---
title: "HW5 - tmc2184"
author: "Talea Cornelius"
date: "11/6/2020"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(RCurl)
library(tidyverse)
```

> Problem 1

>The Washington Post has gathered data on homicides in 50 large U.S. cities and made the data available through a GitHub repository here. You can read their accompanying article here.


```{r load data}
homicide_dat <- read.csv(text = getURL("https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv")) 
names(homicide_dat)
```

>Describe the raw data. Create a city_state variable (e.g. “Baltimore, MD”) and then summarize within cities to obtain the total number of homicides and the number of unsolved homicides (those for which the disposition is “Closed without arrest” or “Open/No arrest”).

```{r tidy data and add summary variables}
homicide_dat = homicide_dat %>%
  janitor::clean_names() %>%
  mutate(state = ifelse(city == "Tulsa", "OK", state)) %>%
  mutate(city_state = paste(city, sep = ", ", str_to_upper(state))) %>%
  group_by(city_state) %>%
  mutate(total_number = n(),
         total_unsolved = sum(disposition != "Closed by arrest")) %>%
  select(city_state, everything())
```

**This dataset comprises details of `r nrow(homicide_dat)` unsolved homicides from 50 major US cities in `r nrow(distinct(select(homicide_dat,state)))` states. There are a total of `r ncol(homicide_dat)-1` variables in the original dataset: `r names(homicide_dat)[1:11]`, and `r names(homicide_dat)[12]`. The maximum number of homicides in a city was `r max(homicide_dat$total_number)`, the minimum was `r min(homicide_dat$total_number)`. The maximum number of unsolved homicides in a city was `r max(homicide_dat$total_unsolved)`, the minimum was `r min(homicide_dat$total_unsolved)`**

>For the city of Baltimore, MD, use the prop.test function to estimate the proportion of homicides that are unsolved; save the output of prop.test as an R object, apply the broom::tidy to this object and pull the estimated proportion and confidence intervals from the resulting tidy dataframe.

```{r prop test for baltimore}
prop_states = function(state) {
  n = mean(state$total_number)
  un = mean(state$total_unsolved)
  prop.test(un, n)
}

baltimore = homicide_dat %>% subset(city_state == "Baltimore, MD") %>%
  prop_states() %>%
  broom::tidy() %>% 
  select(estimate, conf.low, conf.high)

baltimore

```

>Now run prop.test for each of the cities in your dataset, and extract both the proportion of unsolved homicides and the confidence interval for each. Do this within a “tidy” pipeline, making use of purrr::map, purrr::map2, list columns and unnest as necessary to create a tidy dataframe with estimated proportions and CIs for each city.

```{r prop test for each city}

homicide_dat = homicide_dat %>% 
  nest(data = uid:total_unsolved) %>%
  mutate(models = map(data, prop_states),
         tidied = map(models, broom::tidy)) %>%
  unnest(tidied) %>%
  select(city_state, estimate, conf.low, conf.high)

```

>Create a plot that shows the estimates and CIs for each city – check out geom_errorbar for a way to add error bars based on the upper and lower limits. Organize cities according to the proportion of unsolved homicides.

```{r plots for prop tests}

homicide_dat %>% ungroup() %>%
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +  
  geom_point(aes(color = estimate)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high))  +
  theme(text = element_text(size = 15),
        axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1, size = 8),
                                   legend.position = "none") +
  labs(
    title = "Proportion of Unsolved Homicides",
    x = "Location",
    y = "Proportion"
  ) +
  scale_color_gradient(low = "darkblue", high = "red") 

```


>Problem 2

>This zip file contains data from a longitudinal study that included a control arm and an experimental arm. Data for each participant is included in a separate file, and file names include the subject ID and arm.

>Create a tidy dataframe containing data from all participants, including the subject ID, arm, and observations over time:

>Start with a dataframe containing all file names; the list.files function will help
Iterate over file names and read in data for each subject using purrr::map and saving the result as a new variable in the dataframe


```{r read and tidy data}
paths = list.files(path = "./data/", pattern = "*.csv", full.names = T) %>% 
    map_df(~read_csv(.)) %>%
  mutate(file = list.files(path = "./data/"),
         condition = substr(file, 1, 3),
         id = extract_numeric(file)) %>%
  pivot_longer(
    week_1:week_8,
    names_to = "week", 
    values_to = "outcome") %>%
  mutate(week = extract_numeric(week)) %>%
  select(id, condition, week, outcome)

```


>Tidy the result; manipulate file names to include control arm and subject ID, make sure weekly observations are “tidy”, and do any other tidying that’s necessary
Make a spaghetti plot showing observations on each subject over time, and comment on differences between groups.

```{r create plot}
paths %>% 
  ggplot(aes(x = week, y = outcome, group = id, color = condition)) +
    geom_line() + facet_grid(. ~ condition)
```

**The plot shows that subjects in the control (con) and experimental (exp) condition start at around the same level for the outcome at week 0. However, whereas the control condition stays relatively stable, each subject in the experimental condition shows a steady increase in the outcome over time between weeks 0 and 8.**

>Problem 3

>When designing an experiment or analysis, a common question is whether it is likely that a true effect will be detected – put differently, whether a false null hypothesis will be rejected. The probability that a false null hypothesis is rejected is referred to as power, and it depends on several factors, including: the sample size; the effect size; and the error variance. In this problem, you will conduct a simulation to explore power in a simple linear regression.

>First set the following design elements:
Fix n=30
Fix σ=5
Set μ=0. Generate 5000 datasets from the model
x∼Normal[μ,σ]

```{r simulate data}

sim_data = function(mean_val) {tibble(
    x = rnorm(n = 30, mean = mean_val, sd = 5)
  )
}

output_sim = vector("list", 5000)

for (i in 1:5000) {
  output_sim[[i]] = sim_data(0)
}

```

>For each dataset, save μ^ and the p-value arising from a test of H:μ=0 using α=0.05. Hint: to obtain the estimate and p-value, use broom::tidy to clean the output of t.test.

```{r simulate for mean of zero}
tidy_t = function(dat) {
  broom::tidy(t.test(dat))
}

sim_tidy_t = function(mean_val) {
  
  temp_res = vector("list", 5000)

for (i in 1:5000) {
  temp_res[[i]] = tidy_t(sim_data(mean_val))
}
  
  unnest(tibble(temp_res)) %>%
  mutate(mu = mean_val) %>%
  select(mu, estimate, p.value)
  
  
}

mu_0 = sim_tidy_t(0)
mu_0
```


>Repeat the above for μ={1,2,3,4,5,6}, and complete the following:

```{r simulate data for all mu}
output = vector("list", 7)

for (i in 1:7){
  output[[i]] = sim_tidy_t(i-1)
}

sim_results = bind_rows(output)

```

>Make a plot showing the proportion of times the null was rejected (the power of the test) on the y axis and the true value of μ on the x axis. Describe the association between effect size and power.

```{r plot power}

sim_results %>%
  mutate(reject = (p.value < .05)) %>%
  group_by(mu) %>%
  summarise(n_rejected = sum(reject)) %>%
  mutate(proportion_rejected = n_rejected/5000) %>%
  ggplot(aes(x = mu, y = proportion_rejected, color = mu)) +
    geom_point(aes(size = 5)) +
  labs(
    title = "Power and Effect Size, Given N = 30 and SD = 5",
    x = "Mu",
    y = "Proportion Rejected"
  ) +  
  theme_minimal() + 
  theme(legend.position = "none")
```

**This plot is based on simulations of 5,000 datasets (n=30 and sd=5) for 7 different values of mu (0, 1, 2, 3, 4, 5, and 6). The plot shows that, as mu increases, the null hypothesis is rejected more frequently (i.e., as the effect size increases, the power to detect a mean that is not equal to zero also increases).**

>Make a plot showing the average estimate of μ^ on the y axis and the true value of μ on the x axis. Make a second plot (or overlay on the first) the average estimate of μ^ only in samples for which the null was rejected on the y axis and the true value of μ on the x axis. Is the sample average of μ^ across tests for which the null is rejected approximately equal to the true value of μ? Why or why not?

```{r plot estimates}

sim_results %>%
  mutate(reject = (p.value < .05)) %>%
  group_by(mu) %>%
  summarise(mean_all = mean(estimate),
            n_rejected = sum(reject),
            mean_rejected = sum(estimate*reject)/n_rejected) %>%
  pivot_longer(
    c(mean_all, mean_rejected),
    names_to = "estimate_type", 
    values_to = "mean_estimate") %>%
  ggplot(aes(x = mu, y = mean_estimate, group = estimate_type, 
             color = estimate_type)) +
    geom_point() +
  labs(
    title = "True and Estimated Mu",
    x = "True",
    y = "Estimate"
  ) +  
  theme_minimal() + 
  theme(legend.position = "bottom")

```

**The sample average mu across tests is extremely close to the population mu (e.g., across 5,000 repetitions of a random sample where true population mu was equal to 1, the mean estimated sample mu was essentially 1). However, when the null hypothesis was incorrectly rejected, the sample estimate of mu was noticeably different than the true population value (e.g., foa population mu of 1, if the null was rejected, sample mu was estimated to be closer to 2). This makes sense, because random sampling error will sometimes generate mean values that are different from the population mean, and it is more likely that the null hypothesis will be incorrectly rejected when this happens.**




